---
title: "Covid91 vaccine study Final2024"
author: "Melanie Vera"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Question: How effective is the vaccine on males?

## Methods

$H_0$:The vaccine is not effective in preventing COVID19 in males.

$H_A$:The vaccine is effective in preventing COVID19 in males.

Infected ~ Sex
Categorical ~ Categorical
We will use the helper file Cat~Cat

## Graphical
```{r}
Male <- FinalDat[FinalDat$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```

The bar chart shows that males who received the vaccine are less likely to contact covid19.

## Numerical
```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 73.80% of males got infected with covid19, while of the males who received the vaccine only 26.20% contracted covid19. Therefore this indicates that the males who received the vaccine significantly reduced the risk of contracting covid19.

## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-square p-value is 0.00000000000000022. Fisher exact test shows the p-value is 0.00000000000000022 which is lower than 0.05 so we reject the null hypothesis.

## Conclusion

Regarding the males, the bar chart shows you are less likely to contract covid19 if you received the vaccine. The numerical data shows that within the placebo group 73.80% contracted covid19, while the males who received the vaccine significantly reduced the risk of contracting lower than 0.05 so we reject the null hypothesis.

## Females

Question: How effective is the vaccine on females?

##Methods
$H_0$:the vaccine is not effective in preventing covid19 in females.

$H_A$:The vaccine is effective in preventing covid19 in females.

Infected ~ Sex
categorcal ~ categorical
We will use a helper file Cat~Cat

## Graphical
```{r}
Female <- FinalDat[FinalDat$sex=="F",]
```

```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```

The bar chart shows that females that received the vaccine are less likely to contract covid19.

## Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 64.93% of females got infected with covid19, while of the females who received the vaccine only 35.07% contracted covid19. Therefore this indicates that the females who received the vaccin significantly reduced the risk of contracting covid19.

## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test shows the p-value is 0.00000001914 and the Fisher exact test shows the p-value is 0.00000001914 which is lower than 0.05, so we reject the null hypothesis.

## Conclusion

The bar chart shows that females that received the vaccine are less likely o contract covid19. The numerical data shows that within the placebo group 64.93% of females go infected with covid19. Therefore this indicates that the females who recived the vaccine significantly reduced the risk of contracting covid19. The Fisher exact test shows the p-value is 0.00000001914 which is lower than 0.05 so we reject the null hypothesis.

## LGBTQ

Questions: How effective is the vaccine on LGBTQ?

## Methods

$H_0$:The vaccine is not effective in preventing covid 19 in LGBTQ.

$H_A$:The vaccine is effective in preventing covid19 in LGBTQ.

Infected ~ LGBTQ(gay)
Categorical ~ categorical
We will use a helper file Cat~Cat

## Graphial

```{r}
LGBTQ <- FinalDat[FinalDat$LGBTQ=="gay",]
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

The bar chart shows that members of the LGBTQ community are less likely to contract covid19, if they received the vaccine.

## Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 80% of the LGBTQ community members got infected with covid19, while of the LGBTQ community members who recieved the vaccine only 20% contracted covid19.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the chi-square test p-value is 0.0000002266 and the Fisher exact test shows the p-value is 0.0000007269 which is lower than 0.05 so we reject the null hypothesis.

## Conculsion

The bar chart shows that members of the LGBTQ community are less likely to contract covid19 if they received the vaccine. The numerical data shows that within the placebo group 80% of the LGBTQ community members got infected with covid19, while of the LGBTQ community members who received the vaccine only 20% contracted covid19. The Fisher exact test shows the p-value is 0.0000007269 which is lower than 0.05, so we reject the null hypothesis.

## Druggies

Question: How effective is the vaccine on Druggies?

## Methods

$H_0$:The vaccine is not effective in preventing covid19 in Druggies.

$H_A$:The vaccine is effective in preventing covid19 in Druggies.

## Graphical

```{r}
DrugUser <- FinalDat[FinalDat$DrugUser=="yes",]
```


```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

According to the bar chart you are more likely to get covid19 if recieve the vaccine as a drug user.

## Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that that within the placebo group 8.33% of individuals got infected with covid19. While in the vaccine group 91.67% contracted covid19. This indicates that receiving the vaccine as a drug user significantly increases the risk of contracted covid19.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test shows a p-value of 0.0000000001414 data and the Fisher exact test shows that the p-value is 0.000000000003261 which is significantly lower than 0.05 which means that we reject the null hypothesis.

## Conclusion

The bar chart shows that yu are far more likely to contract covi19 if you did receive the vaccine as a DrugUser. The numerical data shows that within the placebo group 91.67% contracted covid19. This indicates that receiving the vaccine as a DrugUser significantly increases the risk of contracting covid19. The inferential data shows a chi-square p-value of 0.0000000001414, and the dtat includes a Fisher exact test showing that the p-value is 0.000000000003261 which is significantly lower than 0.05 which means we reject the null hypothesis. As a DrugUser, you are more likely to contract covid19 if you received the vaccine.

# Overall Results and Conclusions

## Male Conclusion

When trying to figure out the effectiveness of the vaccine in males, the bar chart shows that you are less likely to contact covid19 if you did receive the vaccine. The numerical data shows the results show that within the placebo group 73.80% of individuals got infected with covid19. While in the vaccine group 26.20% contracted covid19. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted covid19. The inferential data shows the chi-square p-value is 0.00000000000000022, and the data also shows th Fisher exact test that shows that the p-value s 0.00000000000000022 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and stimulations show that is as a male you are far less likely to contract covid19 upon receiving the vaccine.

## Female Conclusion

When trying to figure out the effectiveness of the vaccine in females, the bar chart shows that you are less likely to contact covid19 if you did receive the vaccine. The numerical data shows the results show that within the placebo group 64.93% of individuals got infected with covid19. While in the vaccine group 35.07%% contracted covid19. This indicates that receiving the vaccine as a female significantly reduces the risk of contracted covid19. The inferential data shows the chi-square p-value is 0.00000001914, and the data also shows th Fisher exact test that shows that the p-value s 0.00000001246 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and stimulations show that is as a female you are far less likely to contract covid19 upon receiving the vaccine.

## LGBTQ Conclusion
When trying to figure out the effectiveness of the vaccine on members of the LGBTQ comunity, the bar chart shows that you are less likely to contact covid19 if you did receive the vaccine. The numerical data shows the results show that within the placebo group 80% of individuals got infected with covid19. While in the vaccine group 20% contracted covid19. This indicates that receiving the vaccine as a LGBTQ member significantly reduces the risk of contracted covid19. The inferential data shows the chi-square p-value is 0.000002266, and the data also shows th Fisher exact test that shows that the p-value s 0.0000007269 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and simulations show that is as a member of the LGBTQ community you are far less likely to contract covid19 upon receiving the vaccine.

## DrugUser Conclusion

When trying to figure out the effectiveness of the vaccine on people who are DrugUsers, the bar chart shows that you are less likely to contact covid19 if you did receive the vaccine. The numerical data shows the results show that within the placebo group 8.33% of individuals got infected with covid19. While in the vaccine group 91.67% contracted covid19. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted covid19. The inferential data shows the chi-square p-value is 0.0000000001414, and the data also shows th Fisher exact test that shows that the p-value s 0.000000000003261 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and simulations show that is as a DrugUser you are far less likely to contract covid19 upon receiving the vaccine.

## Overall Conclusion

Overall, upon reviewing the data, it is clear that receiving the vaccie will always have the effect on the person receiving it. The data clearly shows that receiving the vaccine is generally effective in preventing the contraction of covid19 if you are male, female, or a member of the LGBTQ community. However, the opposite is true for people who are DrugUsers. In the studies conducted, you are at a much higher risk of contracting covid19 if you are a DrugUser. Presumbly the opposite effect of the vaccine's purpose.